---
title: "02 Redemption"
author: "Laven"
date: "2025-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Set the working directory
```{r}
nile <-scan ("nile.txt")
nile
```
- Check the data type in variable *nile*
- Plot histogram
```{r}
typeof(nile)
plot(nile)
hist(nile)
```
Read data from *nile.txt* and store each data as a "character" type.
```{r}
nile_char <- scan ("nile.txt",what = character())
nile_char
```

Check the data type in *nile_char*

```{r}
typeof(nile_char)
```

Read the plain text data seperated by *";"* for each data, specify the value for argument *sep*
default delimeter "white-space"
```{r}
sunset_sem <- scan("sunset-salvo-amp.txt", what=character(), sep=";")
sunset <- scan("sunset-salvo.txt", what=character())
```

## Reading text data of different format

Reading data with different data types.
Skip the 1st line.
```{r}
running_lst <- scan("running_dat.tsv", what=list("",1.0,1.0,1.0,1.0,1.0,1.0), sep="\t", skip=1)
running_lst
```
### Export Plain Text Files in R

Output *sunset* data with "&" as separator.

```{r}
cat(sunset, file="sunset-salvo-amp.txt", sep="&")
```

## Import Delimiter-Separated Values Files in R

Read in structured data files, into R program as a table (*data frame*), automatically detect the data types for each field/variable.

```{r}
running_tbl <- read.table("running_dat.tsv", header=TRUE, sep="\t")
running_tbl
```

Show the structure of the table/data frame, also to show the data types for each variables.

```{r}
str(running_tbl)
```
### Reading CSV files in R

Read CSV files into R program as data frame.
*read.csv* read comma separated values, with "." as decimal point
*read.csv2* read semicolon separated values, with "," as decimal point
df is data frame
```{r}
running_df <- read.csv("running_dat.csv")
running_df
```

## Export Data Frame in R

Export data frame / table to a delimiter-separated values file.
Example, use "%" as separator, "*" as decimal.

```{r}
write.table(running_df, file="running_star.txt", sep="%", dec="*")
```


## Import XML in R

o	install.packages(“xml2”)
o	library(“xml2”)

Read data from XML into R as a tree hierarchical structure

```{r}
stats_people <- read_xml("statisticians.xml")
stats_people
```

### Export XML Files in R

```{r}
bayes_record <- xml_children(stats_people)[[4]]
xml_remove(bayes_record)
write_xml(stats_people, "statisticians_no_bayes.xml")

```